On the Bit Complexity of Finding Points in Connected Components of a Smooth Real Hypersurface

### 2020-02-07 main.tex
Update main.tex
Section 5

\section{Bit complexity}
Recall Theorem 1.1:   Suppose that $f\in\ZZ[X_1\hdots,X_n]$ is squarefree, satisfies
  $\deg(f) \leq d, \htt(f) \leq s$, and that $V(f) \subset \C^n$ is
  smooth.  Assume also that $f$ is given by a straight-line program
  $\Gamma$ of size $L$ with integer constants of height at most $b.$

  As part of the randomization, the algorithm chooses a set 
  \[
  S := \{1,2,\hdots,2\epsilon^{-1}3n^2(2d)^{3n}\}
  \]
  from which to sample from.

  With probability at least $1-\epsilon$, the output is correct. Otherwise, the algorithm either returns a proper subset of the points or FAIL. In any case, the algorithm uses
  \[
  O^{\sim}(nk(Lb+d^{2n}(s + \log 1/\epsilon+d)(L+d)\log(s + \log 1/\epsilon)))
  \]
  bit operations, with $k \geq 1+\lg n + \lg(1/\epsilon)$. The polynomials in
  the output have degree at most $d^n,$ and height 
  \[
  O^{\sim}((s+\lg 1/\epsilon + d)d^n)
  =O^{\sim}((s+\lg 1/\epsilon)d^{n+1}).
  \]
%
\subsection{The Algorithm}
%
For a matrix $A \in \GL(n,\Q),$ let 
\[
Y_i = X_i^{A} = [b_i] \cdot \bbm X_1\\ \vdots \\ X_n \ebm, 
\]
where $b_i$ is the i-th row of $A^{-1}$.
%
\begin{algorithm}[hbt!]
\KwIn{a straight line program $\Gamma$ that computes $f \in \Q[X_1,\hdots,X_n],$ and $d=\deg(f)$}
\KwOut{$n$ zero-dimensional parameterizations, the union of which includes at least one point in each connected component of $V(f) \cap \R^n$}
\nl Take  $S := \{1,2,\hdots,2\epsilon^{-1}3n^2(2d)^{3n}\}$
 and randomly choose $A \in S^{n^2}$, and  $\sigma_{n-1}=(\sigma_1,\hdots,\sigma_{n-1}) \in S^{n-1}$\;
\nl Compute $f$ and apply the change of variables to obtain $f(A x)=f^A(x)$\;
\nl Use \cite{partialDerComputation} to compute the partial derivatives $\frac{\pa f^A}{\pa X_{i+1}},\hdots,\frac{\pa f^A}{\pa X_n}$\;
    \caption{{Main Algorithm} \label{}}
\nl \For{$i\gets1$ \KwTo $n$}{
\nl Run \cite[Algorithm 2]{SH} $k \geq 1+\lg n + \lg(1/\epsilon)$ times with input (state this in terms of straight line programs?)
$\{Y_1-\sigma_1,\hdots,Y_{i-1}-\sigma_{i-1},f^A,\frac{\pa f^A}{\pa X_{i+1}},\hdots,\frac{\pa f^A}{\pa X_n} \} \textrm{ and } d$\;
\nl Let $\mathscr{Q}_i$ be the highest cardinality zero-dimensional parameterization returned in step 5 \;
} 
\nl \Return $[\mathscr{Q}_1,\hdots,\mathscr{Q}_n]$.
\end{algorithm}

%
\subsection{Proof of Theorem 1.1}
%
\textbf{Bit operation cost:} The following lists the costs for each step of Algorithm 1:
%
\begin{enumerate}
    \item We defined $S = \{1,2,\hdots,2\epsilon^{-1}3n^2(2d)^{3n}\}$ and therefore the height of $a_{i,j} \in S$ is at most 
    \[
    \log 1/{\epsilon} + \log(3n^2(2d)^{2n}) = O^{\sim}(\log 1/{\epsilon}).
    \]
    \item 
    \begin{enumerate}[label=\roman*.]
        \item $O(L)$ operations are required to compute $f$ from its straight line program $\Gamma.$
        \item After applying the change of variables obtaining $f(A x)=f^A(x)$ the height grows at most by a factor of
        \begin{align*}
        \log(n!\cdot \max_{\frac{1\leq i\leq n}{1\leq j\leq n}}\{a_{ij}\}) &= 
        O^{\sim}(n\log n + \log 1/{\epsilon})\\
        &= O^{\sim}(\log 1/{\epsilon}).
        \end{align*}
    \end{enumerate}
    \item After computing the partial derivatives the height grows by at most another factor of $\log d$. Furthermore, preparing the straight line program for the gradient of $f$ does not introduce large integers. The integers in $\grad (f)$ are of the same magnitude as in $\Gamma$.
    \item 
    After applying \cite[Algorithm 2]{SH} $kn$ times, we obtain the
    following upper bound on the total number of bit operations required by the algorithm:
  \[
  O^{\sim}(nk(Lb+d^{2n}(s + \log 1/\epsilon+d)(L+d)\log(s + \log 1/\epsilon)))
  \]
    where it also follows from \cite[Algorithm 2]{SH} that the polynomials in the output have degree at most $d^n,$ and height at most
  \[
  O^{\sim}((s+\lg 1/\epsilon + d)d^n)
  =O^{\sim}((s+\lg 1/\epsilon)d^{n+1}).
  \]
\end{enumerate}
%
\par \noindent 
\textbf{Probability of success:} By Theorem 2.1, there exists $\Delta_i \in \C[\A]$ with $\deg \Delta_i \leq 2n^2(2d)^{3n}.$ By Theorem 2.2, there exists $\Xi_i \subset \C[\A]$ with $\deg \Xi_i \leq d^{2n}.$ Take $\Delta =  \prod_{i=1}^n\Delta_i  \Xi_i $ so that 
%
\begin{align*}
    \deg \Delta &\leq \sum_{i=1}^n (\deg \Delta_i + \deg \Xi_i)\\
    &\leq 2n^2(2d)^{3n} + nd^{2n}\\
    &\leq 3n^2(2d)^{3n}, 
\end{align*}
%
and note that, for $A \in \C^{n \times n}$ that does not cancel $\Delta,$  $\mA$ is then invertible and $f^\mA$ satisfies $\textbf{H}_i$ and $\textbf{H}_i^{'}$ for all $i \in \{1,\hdots,n\}.$

Let $\mathscr{E}$ be the event that, the parameterizations $[\mathscr{Q}_1,\hdots,\mathscr{Q}_n]$ returned in step 7 of Algorithm 1 contain all points and are correct. Then, 
\begin{align*}
\pr[\textrm{success}] = \pr[A \not\in \Delta] \times \pr[\mathscr{E}~|~A \not\in \Delta].  
\end{align*}
By construction of $S$,
\[
\pr[A \not\in \Delta] = 1-\pr[A \in \Delta] \geq
1-\epsilon/2.
\]
Now set $k = \lg(2n/\epsilon)$ so that 
\[
(1-2^{-k})^n = (1 - \epsilon/(2n))^n \geq 1 - \epsilon/2,
\]
by Bernoulli's inequality. Therefore, 
\begin{align*}
\pr[\textrm{success}] &\geq  (1-\epsilon/2)\pr[\mathscr{E}~|~A \not\in \Delta]\\
&\geq (1-\epsilon/2)(1-(11/32)^k)^n\\
& \geq (1-\epsilon/2)(1-2^{-k})^n \\
& \geq (1-\epsilon/2)(1-\epsilon/2)\\
& = 1 - \epsilon + \epsilon^2/4 \geq 1- \epsilon.
\end{align*}
%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-base}
%%
\end{document}

>>>>

\section{Bit complexity}%
\subsection{The Algorithm}
%
For a matrix $A \in \GL(n,\Q),$ let 
\[
Y_i = X_i^{A} = [b_i] \cdot \bbm X_1\\ \vdots \\ X_n \ebm, 
\]
where $b_i$ is the i-th row of $A^{-1}$.
%
\begin{algorithm}[hbt!]
\KwIn{a strait line program $\Gamma$ that computes $f \in \Q[X_1,\hdots,X_n],$ and $d=\deg(f)$}
\KwOut{$n$ zero-dimensional parameterizations, the union of which includes at least one point in each connected component of $V(f) \cap \R^n$}
\nl Take $S := \{1,2,\hdots,3\epsilon^{-1}6n^3(2d)^{2n}\}$ and $T := \{1,2,\hdots,3\epsilon^{-1}nd^{2n}\},$ and randomly choose $A \in S^{n^2}$, and  $\sigma \in T^{n-1}$\;
\nl Apply the change of variables to obtain $f(A x)=f^A(x)$\;
\nl Use \cite{partialDerComputation} to compute the partial derivatives $\frac{\pa f^A}{\pa X_{i+1}},\hdots,\frac{\pa f^A}{\pa X_n}$\;
    \caption{{Main Algorithm} \label{}}
\nl \For{$i\gets1$ \KwTo $n$}{
\nl Run \cite[Algorithm 2]{SH} $k \geq \lg(3n/\epsilon)$ times with a straight-line program computed from 
$\{Y_1-\sigma_1,\hdots,Y_{i-1}-\sigma_{i-1},f^A,\frac{\pa f^A}{\pa X_{i+1}},\hdots,\frac{\pa f^A}{\pa X_n} \}$\;
\nl Let $\mathscr{Q}_i$ be the highest cardinality zero-dimensional parameterization returned in step 5 \;
} 
\nl \Return $[\mathscr{Q}_1,\hdots,\mathscr{Q}_n]$.
\end{algorithm}
%
%
\subsection{Proof of Theorem 1.1}
%
\textbf{Bit operation cost:} The following lists the costs for each step of Algorithm 1:
%
\begin{enumerate}
    \item We defined $S = \{1,2,\hdots,2\epsilon^{-1}3n^2(2d)^{3n}\}$ and therefore the height of $a_{i,j} \in S$ is at most 
    \[
    \log 1/{\epsilon} + \log(3n^2(2d)^{2n}) = O^{\sim}(n\log d + \log 1/{\epsilon}).
    \]
    \item  We can prepare a straight-line program $\Gamma$ for $f^A$ with length at most $d^n$, and 
    by \cite[Lemma 1.2(1)(c)]{SharpEstimatesForTheEffectiveN}, height of at most 
    \[
        O^{\sim}(b + d\log 1 / \epsilon + dn).
    \]
    \item 
    After computing the partial derivatives the height grows by at most another factor of $\log d$. Furthermore, preparing the straight line program for the gradient of $f$ does not introduce large integers. The integers in $\grad (f)$ are of the same magnitude as in $\Gamma$.
    \item 
    After applying \cite[Algorithm 2]{SH} $kn$ times with $k = O(\log n + \log 1 / \epsilon)$, the total boolean cost of the algorithm is
  \[
O^{\sim}(d^{3n+1}(\log1/\epsilon)(b + \log1/\epsilon))
  \]
    where the polynomials in the output have degree at most $d^n,$ and height at most
  \[
  O^{\sim}(d^{n+1}(b + \log 1 / \epsilon)).
  \]
\end{enumerate}
%
\par \noindent 
\textbf{Probability of success:} Let $\Delta_i \in \C[\A]$ be the polynomials from Theorem 2.1. Denote by $\Delta$ the product $\Delta := \prod_{i=1}^n \D_i.$ Note that 
\begin{align}
    \deg \Delta \leq \sum_{i=1}^n \deg \Delta_i \leq 6n^3(2d)^{2n},
\end{align}
where then, if $\mA \in \C^{n \times n}$ does not cancel $\Delta,$ then  $\mA$ is invertible and $f^\mA$ satisfies $\textbf{H}_i$ for all $i \in \{1,\hdots,n\}.$ Now, assuming that $\mA$ is such a matrix, let $\Xi_i\in \C[\A]$ be the polynomials from Theorem 2.2 and denote by $\Xi$ the product $\Xi := \prod_{i=1}^n \Xi_i.$ Note that 
\begin{align}
    \deg \Xi \leq \sum_{i=1}^n \deg \Xi_i \leq nd^{2n},
\end{align}
where then, if $\sigma \in \C^{i-1}$ does not cancel $\Xi$ then $f^{\mA}$ and $\sigma$ satisfy $\textbf{H}_i^{'}$ for all $i \in \{1,\hdots,n\}.$ 
Now, by construction of \[S := \{1,2,\hdots,3\epsilon^{-1}6n^3(2d)^{2n}\}\] and \[T := \{1,2,\hdots,3\epsilon^{-1}nd^{2n}\},\] where $A \in S^{n^2}$ and  $\sigma \in T^{n-1}$ are randomly chosen, we have 
%
\[
\pr[A \in \Delta] \leq  \frac{\deg\Delta}{|S|} = \epsilon/3
\]
and
\[
\pr[\sigma \in \Xi] \leq \deg \frac{\deg\Xi}{|S|} = \epsilon/3.
\]
%
Let $\mathscr{E}$ be the event that, the parameterizations $[\mathscr{Q}_1,\hdots,\mathscr{Q}_n]$ returned in step 7 of Algorithm 1 contain all points and are correct. Then, the probability of success is equal to
\begin{align*}
 \pr[A \not\in \Delta] \times \pr[\sigma \not \in \Xi ~|~A \not\in \Delta] \times \pr[\mathscr{E}~|~A \not\in \Delta \textrm{ and } \sigma \not \in \Xi].
\end{align*}
%
Set $k = \lg(3n/\epsilon)$, with $\epsilon>0,$ so that 
\[
(1-2^{-k})^n = (1 - \epsilon/(3n))^n \geq 1 - \epsilon/3,
\]
by Bernoulli's inequality. Therefore, 
\begin{align*}
\pr[\textrm{success}] &\geq  (1- \epsilon/3)(1- \epsilon/3)\pr[\mathscr{E}~|~A \not\in \Delta \textrm{ and } \sigma \not \in \Xi]\\
&\geq (1- \epsilon/3)(1- \epsilon/3)(1-(11/32)^k)^n\\
& \geq (1- \epsilon/3)(1- \epsilon/3)(1-2^{-k})^n \\
& \geq (1- \epsilon/3)(1- \epsilon/3)(1- \epsilon/3)\\
& \geq 1 - \epsilon.
\end{align*}
%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-base}
%%
\end{document}
